---
title: "Differential Expression Analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params: 
  object: object
  study: study
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: true
    code_folding: hide
    html_notebook: default
---

```{r, "DE-lib", include=FALSE}
require(singleCellTK)
require(DT)
```

```{r "DE-import", include=FALSE}
sce <- params$object
study <- params$study
metas <- metadata(sce)$diffExp[[study]]
useAssay <- metas$useAssay
cond1 <- metas$groupNames[[1]]
cond2 <- metas$groupNames[[2]]
```

# Analysis: `r study`

The differential expression was performed by comparing the feature expression of cells belonging to condition "**`r cond1`**" against the cells belonging to condition "**`r cond2`**". The feature expression matrix used was "**`r useAssay`**".  

In the condition "**`r cond1`**", **`r length(which(metas$select$ix1))`** cells were included, while **`r length(which(metas$select$ix2))`** cells were included for condition **`r cond2`**. 

The method used for performing the differential expression analysis was "**`r metas$method`**". For the more information of the method, please refer to [the help page](#).

## Result Table

```{r "DE-table"}
datatable(metas$result, options = list(pageLength = 10))
```

## Visualization

The following plot shows the linear model regression of the top 36 differentially expressed genes, ranked by FDR values. 

```{r "DE-Vis-regression", results="asis", fig.align="center", warning=FALSE, message=FALSE}
plotDEGRegression(sce, study)
```

The following plot shows the violin plot also for comparing the expression of the top 36 genes between the two conditions. 

```{r "DE-Vis-violin", results="asis", fig.align="center", warning=FALSE, message=FALSE}
plotDEGViolin(sce, study)
```

The following heatmap will be divided by the conditions and the regulation. A gene will be determined as up-regulated in condition **`r cond1`** if its log2FC value was greater than 0.25 and FDR is less than 0.05, and down-regulated in condition **`r cond1`** if the log2FC value is less than -0.25. The gene regulation and condition setting will be annotated by default. In addition, if the condition setting was achieved by using a categorical annotation in the object, the annotation used will also be labeled to the column level of the heatmap. 

```{r "DE-Vis-heatmap", results="asis", fig.align="center", warning=FALSE, message=FALSE}
plotDEGHeatmap(sce, study)
```

# SessionInfo

```{r "DropletQC-session-info"}
    sessionInfo()
```
