---
title: "Differential Expression Analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params: 
  object: object
  study: study
  useReducedDim: useReducedDim
  featureDisplay: featureDisplay
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: no
    theme: cosmo
    code_folding: hide
    self_contained: yes
    mode: selfcontained
    keep_md: no
  pdf_document:
    toc: yes
---

```{r, "DE-lib", include=FALSE}
require(singleCellTK)
require(DT)
require(ggplot2)
```

```{r "DE-import", include=FALSE}
sce <- params$object
study <- params$study
useReducedDim <- params$useReducedDim
metas <- metadata(sce)$diffExp[[study]]
featureDisplay <- params$featureDisplay
useAssay <- metas$useAssay
cond1 <- metas$groupNames[[1]]
cond2 <- metas$groupNames[[2]]
helpURL <- paste0("https://camplab.net/sctk/v", packageVersion("singleCellTK"), "/reference/runDEAnalysis.html")
```

# Analysis: `r study`

The differential expression was performed by comparing the feature expression of cells belonging to condition "**`r cond1`**" against the cells belonging to condition "**`r cond2`**". The relationship between the two selected group is could be indicated via the embedding plot below. The feature expression matrix used was "**`r useAssay`**".  

In the condition "**`r cond1`**", **`r length(which(metas$select$ix1))`** cells were included, while **`r length(which(metas$select$ix2))`** cells were included for condition **`r cond2`**. 

The method used for performing the differential expression analysis was "**`r metas$method`**". For the more information of the method, please refer to [the help page](`r helpURL`).

```{r "UMAP", warning=FALSE, message=FALSE, fig.align="center"}
sce$deClass <- "Not involved"
sce$deClass[metas$select$ix1] <- cond1
sce$deClass[metas$select$ix2] <- cond2
twoColors <- discreteColorPalette(2, palette = "ggplot")
colorValue <- c("grey", twoColors[1], twoColors[2])
names(colorValue) <- c("Not involved", cond1, cond2)
plotSCEDimReduceColData(sce, colorBy = "deClass", reducedDimName = useReducedDim, labelClusters = FALSE) + scale_color_manual(values = colorValue)
```

## Result Table

The following table presents the statistic metrics of the differential expression analysis. The DEGs in the table are pre-filtered by FDR value less than 0.05. 

```{r "DE-table"}
resultTable <- getDEGTopTable(sce, useResult = study, log2fcThreshold = 0, labelBy = featureDisplay)
datatable(resultTable, options = list(pageLength = 10, scollX = 400))
```

>The table above can be reordered by clicking on the column name.

# Visualization

## Heatmap

The following heatmap will be divided by the conditions and the regulation. A gene will be determined as up-regulated in condition **`r cond1`** if its log2FC value was greater than 0.25 and FDR is less than 0.05, and down-regulated in condition **`r cond1`** if the log2FC value is less than -0.25. The gene regulation and condition setting will be annotated by default. In addition, if the condition setting was achieved by using a categorical annotation in the object, the annotation used will also be labeled to the column level of the heatmap. 

```{r "DE-Vis-heatmap", results="asis", fig.align="center", warning=FALSE, message=FALSE, fig.width=9, fig.height=10}
plotDEGHeatmap(sce, study, rowLabel = featureDisplay)
```

## Volcano Plot

The following volcano plot contains dots as all the genes, while the log2FC value on the X axis and -log10(FDR) value on the Y axis. That's to say, the genes farther away from the center of the plot (i.e. at the left or right) show larger fold-change between the two conditions, while the genes closer to the top of the plot show higher significance. The genes with positive log2FC values are up-regulated in the group **`r cond1`**, and those with negative log2FC values are up-regulated in **`r cond2`**. 10 genes with top fold-change (absolute value) are labeled with text.  

```{r "DE-Vis-volcano", results="asis", fig.align="center", warning=FALSE, message=FALSE}
plotDEGVolcano(sce, study, featureDisplay = featureDisplay)
```

## Regression Plot

The following plot shows the linear model regression of the top 36 DEGs, ranked by FDR values. 

```{r "DE-Vis-regression", results="asis", fig.align="center", warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
plotDEGRegression(sce, study, labelBy = featureDisplay)
```

## Violin Plot
 
The following plot shows the violin plot also for comparing the expression of the top 36 DEGs between the two conditions. 

```{r "DE-Vis-violin", results="asis", fig.align="center", warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
plotDEGViolin(sce, study, labelBy = featureDisplay)
```

# SessionInfo

```{r "DE-session-info"}
sessionInfo()
```
