---
title: "Find Marker"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params: 
  object: object
  study: study
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: true
    code_folding: hide
    html_notebook: default
---

```{r, "FM-lib", include=FALSE}
require(singleCellTK)
require(DT)
require(knitr)
```

```{r "FM-import", include=FALSE}
sce <- params$object
fmTable <- metadata(sce)$findMarker
useAssay <- attr(fmTable, "useAssay")
method <- attr(fmTable, "method")
clusterName <- names(fmTable)[5]
nCluster <- length(unique(as.vector(sce[[clusterName]])))
uniqClusters <- sort(unique(as.vector(sce[[clusterName]])))
fmParams <- attr(fmTable, "params")
log2fcThreshold <- fmParams$log2fcThreshold
fdrThreshold <- fmParams$ fdrThreshold
minClustExprPerc <- fmParams$minClustExprPerc
maxCtrlExprPerc <- fmParams$maxCtrlExprPerc
minMeanExpr <- fmParams$minMeanExpr
```

# Analysis Setting

## Find Marker basing on: `r clusterName`

The marker finding was accomplished by iteratively performing one-against-others differential expression analysis for each cluster. There are **`r nCluster` clusters** found in the given SingleCellExperiment object. Here is the summary of the number of cells in each cluster:

```{r, "FM-clusterSummary"}
sum.table <- t(as.data.frame(table(sce$scranSNN_cluster)))
rownames(sum.table) <- c("cluster", "cell number")
datatable(sum.table, 
          options = list(
            headerCallback = JS(
              "function(thead, data, start, end, display){",
              "  $(thead).remove();",
              "}"),
            dom = 't')
)
```


## Method: `r method`

The method used for performing the differential expression analysis was "**`r method`**". For the more information of the method, please refer to [the help page](#).

## Parameters

After the computation, we further filtered the result to select the differentially expressed genes with high significance and high fold change. Specifically, the parameters used are:

```{r "FM-params"}
p <- data.frame(parameter = c("Min. logFC", 
                              "Max. FDR", 
                              "Min. Cluster Expression %", 
                              "Max. Control Expression %", 
                              "Min. Mean Expression"),
                value = c(log2fcThreshold,
                          fdrThreshold,
                          minClustExprPerc,
                          maxCtrlExprPerc,
                          minMeanExpr))
datatable(p, options = list(dom = "t"))
```

When we identify a gene to be the marker of the cluster of interests:  

- **"Min. logFC"** - The log fold change of this gene, by comparing the cluster of interest against all other clusters, must be higher than this threshold;
- **"Max. FDR"** - The FDR value calculated, by comparing the cluster of interest against all other clusters, must not exceed this threshold; 
- **"Min. Cluster Expression %"** - The gene must be expressed in at least this fraction of cells in the cluster of interests;
- **"Max. Control Expression %"** - The gene must be expressed in less than this fraction of cells in all other clusters;
- **"Min. Mean Expression"** - The mean expression of the gene in the cluster of interests must be higher than this threshold (in the specific assay used for analysis, not necessarily the raw count matrix).

# Result

We report maximum 10 top markers for each cluster, ranked by the FDR value. 

## Marker Table {.tabset}

```{r FM-iterResTable, echo=FALSE, results="asis"}
out = NULL
for (i in uniqClusters) {
  fmTable.cluster <- fmTable[fmTable[[clusterName]] == i,]
  if (nrow(fmTable.cluster) >= 10) {
    fmTable.cluster.top <- fmTable.cluster[1:10,]
  } else {
    fmTable.cluster.top <- fmTable.cluster
  }
  res <- knit_child('FM_child_cluster.Rmd', quiet=TRUE)
  cat(res, sep = '\n')
}
```

## Heatmap

In the marker heatmap, each column stands for a cell in the given SingleCellExperiment object. The cell level color bar annotates the cluster assignment. Each row is a marker gene, and the gene level color bar assigns the cluster correspondence. The cell cluster is ordered by the cluster population size.  

It happens sometimes that a gene is identified as the marker for more than one cluster. In this scenario, we finally assign it to the cluster where it has a higher "logFC" value when plotting.  

```{r FM-heatmap, message=FALSE, warning=FALSE, out.width = '100%'}
plotMarkerDiffExp(sce, 
                  log2fcThreshold = 0.5, 
                  minClustExprPerc = 0.7, 
                  maxCtrlExprPerc = 0.4, 
                  minMeanExpr = 1)
```
