---
title: "SCTK Cell QC Report"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params: 
  object: object
  subTitle: subTitle
  studyDesign: studyDesign
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: true
    collapsed: true
    code_folding: hide
    html_notebook: default
---

<style>
body {
text-align: justify}
</style>

```{r "CellQC-lib", warning=FALSE, include=FALSE}
require(singleCellTK)
require(ggplot2)
require(umap)
require(dplyr)

preprocessMeta <- function(meta, ignore=c("geneSetCollection", "sample")){
  meta <- meta[!names(meta) %in% ignore]
  lapply(meta, function(y){
    while(is.list(y)) {
      if (length(y) != 0){
       y <- y[[1]]  
      } else {
       y <- "NULL"
      }
    }
      
    if(is.vector(y)) {
      y <- paste(y,collapse=' ')
    }
      
    if(is.null(y)) {
      y <- "NULL"
    }
      
    return(y)
  })
}

formatPlotList <- function(plotlist, samples) {
  if (length(samples) > 1) {
    return(plotlist)
  }
  
  plotlist <- list(Sample = list(plotlist))
  names(plotlist$Sample) <- samples
  return(plotlist)
}

```

```{r, "CellQC-import", eval = TRUE, include=FALSE}
sce.qc <- params$object
subTitle <- params$subTitle
studyDesign <- params$studyDesign
samples <- unique(colData(sce.qc)$sample)

reduceDims<- grep('UMAP', reducedDimNames(sce.qc), value = TRUE)
if (length(reduceDims) != 0) {
  reduceDimUsed <- reduceDims[1]
} else {
  reduceDims<- grep('TSNE', reducedDimNames(sce.qc), value = TRUE)
  if (length(reduceDims) == 0) {
    stop("No reduced dimension are available for QC visualization!")
  } else {
    reduceDimUsed <- reduceDims[1]
  }
}
```

---
subtitle: "`r subTitle`"
---
`r studyDesign`
<br />


# Introduction
Comprehensive quality control (QC) of single-cell RNA-seq data was performed with the [**singleCellTK**](https://github.com/compbiomed/singleCellTK/tree/sctk_qc_1.7.6) package. This report contains information about each QC tool and visualization of the QC metrics for each sample. For more information on running this pipeline and performing quality control, see the [**documentation**](https://github.com/compbiomed/singleCellTK/blob/devel/exec/SCTK_runQC_Documentation2.md). If you use the singleCellTK package for quality control, please include a [**reference**](https://www.biorxiv.org/content/10.1101/329755v1.article-info) in your publication.
<br />
<br />

```{r, include=FALSE, warning=FALSE, message=FALSE}
description_runPerCellQC <- descriptionRunPerCellQC()
```

```{r, "General quality control metrics", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
i="General quality control metrics"
cat(paste0('# ', i, ' \n'))
cat("\n")

perCellData <- c("sum", "detected", "percent_top_50")
skipCellData <- any(!perCellData %in% names(colData(sce.qc)))

if (skipCellData) {
  cat("runPerCellQC did not complete successfully. The section of General quality control metrics is skipped")
  plotsQCMetrics <- NULL
} else {
  plotsQCMetrics<- tryCatch(
    {plotRunPerCellQCResults(sce.qc, 
                             sample = colData(sce.qc)$sample, 
                             combinePlot="none")},
    error = function(e) {
      cat("runPerCellQC did not complete successfully. The section of General quality control metrics is skipped")
      skipCellData <- TRUE
      return(NA)
    }
  )
}

if (!skipCellData) {
  
  if (length(samples) == 1) {
    plotsQCMetrics <- list(Violin=plotsQCMetrics)
  }
  cat(description_runPerCellQC$introduction)
  cat(description_runPerCellQC$runPerCellQC)
  cat(description_runPerCellQC$plotRunPerCellQCResults)
  cat(description_runPerCellQC$output)
  cat(description_runPerCellQC$sum)
  cat(description_runPerCellQC$detected)
  cat(description_runPerCellQC$percentTop)
  cat(description_runPerCellQC$subsets)
}
```


# {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r results="asis", eavl=(!skipCellData), fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
cat("\n")
cat(paste0('## Sum {.tabset .tabset-fade} \n\n'))
plotsQCMetrics$Violin$Sum
cat("\n\n")
cat(paste0('## Total {.tabset .tabset-fade} \n\n'))
plotsQCMetrics$Violin$Detected
cat("\n\n")
cat(paste0('## Subsets mito sum {.tabset .tabset-fade} \n\n'))
plotsQCMetrics$Violin$Gene_Subset_Sum
cat("\n\n")
cat(paste0('## Subsets mito detected {.tabset .tabset-fade} \n\n'))
plotsQCMetrics$Violin$Gene_Subset_Features
cat("\n\n")
cat(paste0('## Subsets mito percent {.tabset .tabset-fade} \n\n'))
plotsQCMetrics$Violin$Gene_Subset_Top50_Percent
cat("\n\n")
cat(paste0('## Session info {.tabset .tabset-fade} \n\n'))

runPerCellMeta <- sce.qc@metadata$scater
x <- preprocessMeta(runPerCellMeta[[samples[1]]]$addPerCellQC)
t(as.data.frame(x)) %>%
  knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
  kableExtra::scroll_box(width = "80%")

cat("\n")
cat(description_runPerCellQC$parameter)
```
`r cat("\n\n")`



# Doublet Detection
```{r, include=FALSE, warning=FALSE, message=FALSE}
description_Scrublet <- descriptionScrublet()
```

```{r, "Scrublet", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo = FALSE}
i="Scrublet"
cat(paste0('## ', i, ' \n'))

scrubletData <- c("scrublet_score", "scrublet_call")
skipScrublet <- any(!scrubletData %in% names(colData(sce.qc)))

if (skipScrublet) {
  cat("runScrublet did not complete successfully. The section of Scrublet is skipped")
  plotsScrublet <- NULL
} else {
    plotsScrublet<- tryCatch(
      {plotScrubletResults(sce.qc, 
                           reducedDimName=reduceDimUsed,
                           sample = colData(sce.qc)$sample,
                           combinePlot = "none")},
    error = function(e) {
      cat("runScrublet did not complete successfully. The section of Scrublet is skipped")
      skipScrublet <- TRUE
      return(NA)
    }
  )
}

if (!skipScrublet) {
  plotsScrublet <- formatPlotList(plotsScrublet, samples)
  cat(description_Scrublet$introduction)
  cat(description_Scrublet$runScrublet)
  cat(description_Scrublet$plotScrubletResults)
  cat(description_Scrublet$output)
}
```
<br />

## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}

```{r echo=FALSE, eval=(!skipScrublet), results="asis", fig.align="center"}
cat("\n")
for (sample in samples) {
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### Scrublet Doublet Score {.tabset .tabset-fade} \n")
  print(plotsScrublet$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n")
  print(plotsScrublet$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n ")
  print(plotsScrublet$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Scrublet Doublet Assignment {.tabset .tabset-fade} \n\n")
  print(plotsScrublet$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")  
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  runScrubletMeta <- sce.qc@metadata$runScrublet
  # if(length(samples) == 1) {
  #   runScrubletMeta <- list(runScrubletMeta)
  #   names(runScrubletMeta) <- sample
  # }
    
  x <- preprocessMeta(runScrubletMeta[[sample]])
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  cat(description_Scrublet$additionalParam)
  cat("\n\n")
}
```
`r cat("\n")` 




```{r, include=FALSE, warning=FALSE, message=FALSE}
description_DoubletFinder <- descriptionDoubletFinder()
```

```{r, "DoubletFinder-0", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
i="DoubletFinder"
cat(paste0('## ', i, ' \n'))


doubletFinderData <- grep("doubletFinder_doublet", names(colData(sce.qc)), value=TRUE)
skipFinder <- length(doubletFinderData)==0  

if (skipFinder) {
  cat("runDoubletFinder did not complete successfully. The section of DoubletFinder is skipped")
  plotDoubletFinder <- NULL
} else {
  plotDoubletFinder<- tryCatch(
    {plotDoubletFinderResults(inSCE = sce.qc, 
                              reducedDimName=reduceDimUsed,
                              sample = colData(sce.qc)[['sample']],
                              combinePlot = "none")},
    error = function(e) {
      cat("runDoubletFinder did not complete successfully. The section of DoubletFinder is skipped")
      skipFinder <- TRUE
      return(NA)
    }
  )
}

if (!skipFinder) {
  plotDoubletFinder <- formatPlotList(plotDoubletFinder, samples)
  resolutions <- grep("doubletFinder_doublet_score_resolution_", colnames(colData(sce.qc)), value = TRUE)
  resolutions <- gsub("doubletFinder_doublet_score_resolution_", "", resolutions)
  
  cat(description_DoubletFinder$introduction)
  cat(description_DoubletFinder$runDoubletFinder)
  cat(description_DoubletFinder$plotDoubletFinderResults)
  cat(description_DoubletFinder$output)
}
```
<br />


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE, eavl=!(skipFinder), results="asis", fig.align="center"}
cat("\n")
for (sample in samples) {
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n")
  for (resolution in resolutions) {
    cat(paste0("#### Resolution: ", resolution, " {.tabset .tabset-fade} \n"))
    cat("\n")
    
    cat("##### DoubletFinder Doublet Score {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("Scatter_Score_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### Density of Doublet Score {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("Density_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### Violin of Doublet Score {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("violin_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### DoubletFinder Doublet Assignment {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("Scatter_Call_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### Parameters {.tabset .tabset-fade} \n\n")
    DoubletFinderMeta <- sce.qc@metadata$runDoubletFinder
    # if(length(samples) == 1) {
    #   DoubletFinderMeta <- list(DoubletFinderMeta)
    #   names(DoubletFinderMeta) <- sample
    # }
    
    x <- preprocessMeta(DoubletFinderMeta[[sample]])
    cat(t(as.data.frame(x)) %>%
      knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
      kableExtra::scroll_box(width = "80%"))
    
    cat(description_DoubletFinder$seuratRes)
    cat(description_DoubletFinder$seuratNfeatures)
    cat(description_DoubletFinder$seuratPCs)
    cat(description_DoubletFinder$seuratFormationRate)
    cat("\n\n")
  }
}
```
`r cat("\n")`








```{r, include=FALSE, warning=FALSE, message=FALSE}
description_DoubletCells<- descriptionDoubletCells()
```

```{r, "DoubletCells", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
i="DoubletCells"
cat(paste0('## ', i, ' \n'))

doubletCellData <- c("scran_doubletCells_score_log10")
skipDoubletCell <- any(!doubletCellData %in% names(colData(sce.qc)))

if (skipDoubletCell) {
  cat("runDoubletCells did not complete successfully. The section of DoubletCells is skipped")
  plotDoubletCells <- NULL
} else {
  plotDoubletCells<- tryCatch(
    {plotDoubletCellsResults(inSCE = sce.qc, 
                             reducedDimName=reduceDimUsed,
                             sample = colData(sce.qc)[['sample']],
                             combinePlot = "none")},
    error = function(e) {
      cat("runDoubletCells did not complete successfully. The section of DoubletCells is skipped")
      skipDoubletCell <- TRUE
      return(NA)
    }
  )
}

if(!skipDoubletCell) {
  plotDoubletCells <- formatPlotList(plotDoubletCells, samples)
  cat(description_DoubletCells$introduction)
  cat(description_DoubletCells$runDoubletCells)
  cat(description_DoubletCells$plotDoubletCellsResults)
  cat(description_DoubletCells$output)
}
```
<br />


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE, eval=!(skipDoubletCell), results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### DoubletCells Doublet Score {.tabset .tabset-fade} \n\n ")
  print(plotDoubletCells$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  print(plotDoubletCells$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  print(plotDoubletCells$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  DoubletCellMeta <- sce.qc@metadata$runDoubletCell
  # if(length(samples) == 1) {
  #   DoubletCellMeta <- list(DoubletCellMeta)
  #   names(DoubletCellMeta) <- sample
  # }
  
  x <- preprocessMeta(DoubletCellMeta[[sample]])
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  cat(description_DoubletCells$parameter)
  cat("\n\n")
}
```
`r cat("\n")`




```{r, include=FALSE, warning=FALSE, message=FALSE}
descriprion_CxdsResults<- descriptionCXDS()
```

```{r, "Cxds", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
i="Cxds"
cat(paste0('## ', i, ' \n'))

cvdsData <- c("scds_cxds_score", "scds_cxds_call")
skipCxds <- any(!cvdsData %in% names(colData(sce.qc)))

if (skipCxds) {
  cat("runCxds did not complete successfully. The section of Cxds is skipped")
  plotCxds <- NULL
} else {
  plotCxds<- tryCatch(
    {plotCxdsResults(inSCE = sce.qc, 
                              reducedDimName=reduceDimUsed,
                              sample = colData(sce.qc)[['sample']],
                              combinePlot = "none")},
    error = function(e) {
      cat("runCxds did not complete successfully. The section of Cxds is skipped")
      skipCxds <- TRUE
      return(NA)
    }
  )
}

if (!skipCxds) {
  plotCxds <- formatPlotList(plotCxds, samples)
  cat(descriprion_CxdsResults$introduction)
  cat(descriprion_CxdsResults$runCxds)
  cat(descriprion_CxdsResults$plotCxdsResults)
  cat(descriprion_CxdsResults$output)
}
```
<br />



## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE,eval=(!skipCxds), results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### Cxds Doublet Score {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Cxds Doublet Assignment {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  CxdsMeta <- sce.qc@metadata$runCxds
  # if(length(samples) == 1) {
  #   CxdsMeta <- list(CxdsMeta)
  #   names(CxdsMeta) <- sample
  # }
  
  x <- preprocessMeta(CxdsMeta[[sample]])
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  
  cat(descriprion_CxdsResults$nTop)
  cat(descriprion_CxdsResults$binThresh)
  cat(descriprion_CxdsResults$verb)
  cat(descriprion_CxdsResults$retRes)
  cat(descriprion_CxdsResults$estNdbl)
  cat("\n\n")
}
```
`r cat("\n")`





```{r, include=FALSE, warning=FALSE, message=FALSE}
descriprion_BcdsResults<- descriptionBCDS()
```

```{r, "Bcds", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
i="Bcds"
cat(paste0('## ', i, ' \n'))

bcdsData <- c("scds_bcds_score", "scds_bcds_call")
skipBcds <- any(!bcdsData %in% names(colData(sce.qc)))

if (skipBcds) {
  cat("runBcds did not complete successfully. The section of Bcds is skipped")
  plotBcds <- NULL
} else {
  plotBcds<- tryCatch(
    {plotBcdsResults(inSCE = sce.qc, 
                              reducedDimName=reduceDimUsed,
                              sample = colData(sce.qc)[['sample']],
                              combinePlot = "none")},
    error = function(e) {
      cat("runBcds did not complete successfully. The section of Bcds is skipped")
      skipBcds <- TRUE
      return(NA)
    }
  )
}

if (!skipBcds) {
  plotBcds <- formatPlotList(plotBcds, samples)
  cat(descriprion_BcdsResults$introduction)
  cat(descriprion_BcdsResults$runBcds)
  cat(descriprion_BcdsResults$plotBcdsResults)
  cat(descriprion_BcdsResults$output)
}
```
<br />


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE, eval=(!skipBcds), results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### Bcds Doublet Score {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Bcds Doublet Assignment {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  BcdsMeta <- sce.qc@metadata$runBcds
  # if(length(samples) == 1) {
  #   BcdsMeta <- list(BcdsMeta)
  #   names(BcdsMeta) <- sample
  # }
  
  x <- preprocessMeta(BcdsMeta[[sample]])
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  
  cat(descriprion_BcdsResults$nTop)
  cat(descriprion_BcdsResults$srat)
  cat(descriprion_BcdsResults$nmax)
  cat(descriprion_BcdsResults$varImp)
  cat("\n\n")
}
```
`r cat("\n")`






```{r, include=FALSE, warning=FALSE, message=FALSE}
description_ScdsHybrid<- descriptionScdsHybrid()
```


```{r, "ScdsHybrid", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
i="ScdsHybrid"
cat(paste0('## ', i, ' \n'))

hybridData <- c("scds_hybrid_score", "scds_hybrid_call")
skipScdsHybrid <- any(!hybridData %in% names(colData(sce.qc)))

if (skipScdsHybrid) {
  cat("runCxdsBcdsHybrid did not complete successfully. The section of ScdsHybrid is skipped")
  plotScdsHybrid <- NULL
} else {
  plotScdsHybrid<- tryCatch(
    {plotScdsHybridResults(inSCE = sce.qc, 
                              reducedDimName=reduceDimUsed,
                              sample = colData(sce.qc)[['sample']],
                              combinePlot = "none")},
    error = function(e) {
      cat("runCxdsBcdsHybrid did not complete successfully. The section of ScdsHybrid is skipped")
      skipScdsHybrid <- TRUE
      return(NA)
    }
  )
}

if (!skipScdsHybrid) {
  plotScdsHybrid <- formatPlotList(plotScdsHybrid, samples)
  cat(description_ScdsHybrid$introduction)
  cat(description_ScdsHybrid$runCxdsBcdsHybrid)
  cat(description_ScdsHybrid$plotScdsHybridResults)
  cat(description_ScdsHybrid$output)
}
```
<br />


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE, eval=(!skipScdsHybrid), results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### ScdsHybrid Doublet Score {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### ScdsHybrid Doublet Assignment {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  HybridMeta <- sce.qc@metadata$runCxdsBcdsHybrid
  # if(length(samples) == 1) {
  #   HybridMeta <- list(HybridMeta)
  #   names(HybridMeta) <- sample
  # }
  
  x <- preprocessMeta(HybridMeta[[sample]])
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  
  cat(description_ScdsHybrid$parameters)
  cat("\n\n")
}
```
`r cat("\n")`




# Ambient RNA Detection
```{r, include=FALSE, warning=FALSE, message=FALSE}
description_DecontX<- suppressWarnings(descriptionDecontX())
```

```{r, "DecontX", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
i="DecontX"
cat(paste0('## ', i, ' \n'))

decontXData <- c("decontX_contamination", "decontX_clusters")
skipDecontX <- any(!decontXData %in% names(colData(sce.qc)))

if (skipDecontX) {
  cat("runDecontX did not complete successfully. The section of DecontX is skipped")
  plotDecontX <- NULL
} else {
  plotDecontX<- tryCatch(
    {plotDecontXResults(inSCE = sce.qc, 
                              reducedDimName=reduceDimUsed,
                              sample = colData(sce.qc)[['sample']],
                              combinePlot = "none")},
    error = function(e) {
      cat("runDecontX did not complete successfully. The section of DecontX is skipped")
      skipDecontX <- TRUE
      return(NA)
    }
  )
}

if (!skipDecontX) {
  plotDecontX <- formatPlotList(plotDecontX, samples)
  cat(description_DecontX$introduction)
  cat(description_DecontX$runDecontX)
  cat(description_DecontX$plotDecontXResults)
  cat(description_DecontX$output)
  cat(description_DecontX$contamination)
  cat(description_DecontX$clustering)
}
```
<br />


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE, eval=(!skipDecontX), results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### DecontX Contamination Score {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["scatter_decontXContamination"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["density_decontXContamination"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["violin_decontXContamination"]])
  cat("\n\n")
  
  cat("#### DecontX Clusters {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["scatter_decontXClusters"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  DecontXMeta <- sce.qc@metadata$runDecontX
  # if(length(samples) == 1) {
  #   DecontXMeta <- list(DecontXMeta)
  #   names(DecontXMeta) <- sample
  # }
  
  x <- preprocessMeta(DecontXMeta[[sample]])
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  cat("\n\n")
}
```
`r cat("\n")`

# Session Info
```{r "CellQC-session-info", echo=FALSE}
sessionInfo()
```




