---
title: "SCTK Cell QC Report"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params: 
  object: object
  subTitle: subTitle
  studyDesign: studyDesign
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: true
    collapsed: true
    code_folding: hide
    html_notebook: default
---

<style>
body {
text-align: justify}
</style>

```{r "CellQC-lib", warning=FALSE, include=FALSE}
require(singleCellTK)
require(ggplot2)
require(umap)
require(dplyr)

preprocessMeta <- function(meta, ignore=c("geneSetCollection", "sample")){
  meta <- meta[!names(meta) %in% ignore]
  lapply(meta, function(y){
    while(is.list(y)) {
      if (length(y) != 0){
       y <- y[[1]]  
      } else {
       y <- "NULL"
      }
    }
      
    if(is.vector(y)) {
      y <- paste(y,collapse=' ')
    }
      
    if(is.null(y)) {
      y <- "NULL"
    }
      
    return(y)
  })
}

formatPlotList <- function(plotlist, samples) {
  if (length(samples) > 1) {
    return(plotlist)
  }
  
  plotlist <- list(Sample = list(plotlist))
  names(plotlist$Sample) <- samples
  return(plotlist)
}

```

```{r, "CellQC-import", eval = TRUE, include=FALSE}
sce.qc <- params$object
subTitle <- params$subTitle
studyDesign <- params$studyDesign
samples <- unique(colData(sce.qc)$sample)

reduceDims<- grep('UMAP', reducedDimNames(sce.qc), value = TRUE)
if (length(reduceDims) != 0) {
  reduceDimUsed <- reduceDims[1]
} else {
  reduceDims<- grep('TSNE', reducedDimNames(sce.qc), value = TRUE)
  if (length(reduceDims) == 0) {
    stop("No reduced dimension are available for QC visualization!")
  } else {
    reduceDimUsed <- reduceDims[1]
  }
}
```

---
subtitle: "`r subTitle`"
---
`r studyDesign`
<br />


# Introduction
Comprehensive quality control (QC) of single-cell RNA-seq data was performed with the [**singleCellTK**](https://github.com/compbiomed/singleCellTK/tree/sctk_qc_1.7.6) package. This report contains information about each QC tool and visualization of the QC metrics for each sample. For more information on running this pipeline and performing quality control, see the [**documentation**](https://github.com/compbiomed/singleCellTK/blob/devel/exec/SCTK_runQC_Documentation2.md). If you use the singleCellTK package for quality control, please include a [**reference**](https://www.biorxiv.org/content/10.1101/329755v1.article-info) in your publication.
<br />
<br />

```{r, include=FALSE, warning=FALSE, message=FALSE}
description_runPerCellQC <- descriptionRunPerCellQC()
```

```{r, "General quality control metrics", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
plotsQCMetrics<-plotRunPerCellQCResults(sce.qc, 
                                        sample = colData(sce.qc)$sample, 
                                        combinePlot="none")
plotsQCMetrics <- formatPlotList(plotsQCMetrics, samples)
i="General quality control metrics"
cat(paste0('# ', i, ' \n'))
```
<br />
`r description_runPerCellQC$introduction`
`r description_runPerCellQC$runPerCellQC`
`r description_runPerCellQC$plotRunPerCellQCResults`
`r description_runPerCellQC$output`
`r description_runPerCellQC$sum`
`r description_runPerCellQC$detected`
`r description_runPerCellQC$percentTop`
`r description_runPerCellQC$subsets`

# {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}

```{r results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
cat("\n")
for (sample in samples) {
  cat(paste0("## ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat(paste0('### Sum {.tabset .tabset-fade} \n\n'))
  print(plotsQCMetrics$Sample[[sample]][["Sum"]])
  cat("\n\n")
  
  cat(paste0('### Total {.tabset .tabset-fade} \n\n'))
  print(plotsQCMetrics$Sample[[sample]][["Detected"]])
  cat("\n\n")
  
  cat(paste0('### Subsets mito sum {.tabset .tabset-fade} \n\n'))
  print(plotsQCMetrics$Sample[[sample]][["Gene_Subset_Sum"]])
  cat("\n\n")
  
  cat(paste0('### Subsets mito detected {.tabset .tabset-fade} \n\n'))
  print(plotsQCMetrics$Sample[[sample]][["Gene_Subset_Features"]])
  cat("\n\n")
  
  cat(paste0('### Subsets mito percent {.tabset .tabset-fade} \n\n'))
  print(plotsQCMetrics$Sample[[sample]][["Gene_Subset_Top50_Percent"]])
  cat("\n\n")

  cat("### Parameters {.tabset .tabset-fade} \n\n")
  runPerCellMeta <- sce.qc@metadata$scater
  # if(length(samples) == 1) {
  #   runScrubletMeta <- list(runScrubletMeta)
  #   names(runScrubletMeta) <- sample
  # }
    
  x <- preprocessMeta(runPerCellMeta$addPerCellQC)
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  cat("\n\n")
}
```
`r cat("\n")`
`r description_runPerCellQC$parameter`
`r cat("\n\n")`



# Doublet Detection
```{r, include=FALSE, warning=FALSE, message=FALSE}
description_Scrublet <- descriptionScrublet()
```

```{r, "Scrublet", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo = FALSE}
plotsScrublet<- suppressWarnings(plotScrubletResults(sce.qc, 
                                                     reducedDimName=reduceDimUsed,
                                                     sample = colData(sce.qc)$sample,
                                                     combinePlot = "none"))
plotsScrublet <- formatPlotList(plotsScrublet, samples)
i="Scrublet"
cat(paste0('## ', i, ' \n'))
```
<br />
`r description_Scrublet$introduction`
`r description_Scrublet$runScrublet`
`r description_Scrublet$plotScrubletResults`
`r description_Scrublet$output`

## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}

```{r echo=FALSE, results="asis", fig.align="center"}
cat("\n")
for (sample in samples) {
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### Scrublet Doublet Score {.tabset .tabset-fade} \n")
  print(plotsScrublet$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n")
  print(plotsScrublet$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n ")
  print(plotsScrublet$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Scrublet Doublet Assignment {.tabset .tabset-fade} \n\n")
  print(plotsScrublet$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")  
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  runScrubletMeta <- sce.qc@metadata$runScrublet
  # if(length(samples) == 1) {
  #   runScrubletMeta <- list(runScrubletMeta)
  #   names(runScrubletMeta) <- sample
  # }
    
  x <- preprocessMeta(runScrubletMeta)
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  cat(description_Scrublet$additionalParam)
  cat("\n\n")
}
```
`r cat("\n")` 






```{r, include=FALSE, warning=FALSE, message=FALSE}
description_DoubletFinder <- descriptionDoubletFinder()
```

```{r, "DoubletFinder-0", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
plotDoubletFinder <- suppressWarnings(plotDoubletFinderResults(inSCE = sce.qc, 
                                                              reducedDimName=reduceDimUsed,
                                                              sample = colData(sce.qc)[['sample']],
                                                              combinePlot = "none"))
plotDoubletFinder <- formatPlotList(plotDoubletFinder, samples)

resolutions <- grep("doubletFinder_doublet_score_resolution_", colnames(colData(sce.qc)), value = TRUE)
resolutions <- gsub("doubletFinder_doublet_score_resolution_", "", resolutions)

i="DoubletFinder"
cat(paste0('## ', i, ' \n'))
```
<br />
`r description_DoubletFinder$introduction`
`r description_DoubletFinder$runDoubletFinder`
`r description_DoubletFinder$plotDoubletFinderResults`
`r description_DoubletFinder$output`


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE,results="asis", fig.align="center"}
cat("\n")
for (sample in samples) {
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n")
  for (resolution in resolutions) {
    cat(paste0("#### Resolution: ", resolution, " {.tabset .tabset-fade} \n"))
    cat("\n")
    
    cat("##### DoubletFinder Doublet Score {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("Scatter_Score_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### Density of Doublet Score {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("Density_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### Violin of Doublet Score {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("violin_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### DoubletFinder Doublet Assignment {.tabset .tabset-fade} \n\n")
    print(plotDoubletFinder$Sample[[sample]][[paste0("Scatter_Call_resolution_", resolution)]])
    cat("\n\n")
    
    cat("##### Parameters {.tabset .tabset-fade} \n\n")
    DoubletFinderMeta <- sce.qc@metadata$runDoubletFinder
    # if(length(samples) == 1) {
    #   DoubletFinderMeta <- list(DoubletFinderMeta)
    #   names(DoubletFinderMeta) <- sample
    # }
    
    x <- preprocessMeta(DoubletFinderMeta)
    cat(t(as.data.frame(x)) %>%
      knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
      kableExtra::scroll_box(width = "80%"))
    
    cat(description_DoubletFinder$seuratRes)
    cat(description_DoubletFinder$seuratNfeatures)
    cat(description_DoubletFinder$seuratPCs)
    cat(description_DoubletFinder$seuratFormationRate)
    cat("\n\n")
  }
}

```
`r cat("\n")`








```{r, include=FALSE, warning=FALSE, message=FALSE}
description_DoubletCells<- descriptionDoubletCells()
```

```{r, "DoubletCells", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
plotDoubletCells<- suppressWarnings(plotDoubletCellsResults(inSCE = sce.qc, 
                                                            reducedDimName=reduceDimUsed,
                                                            sample = colData(sce.qc)[['sample']],
                                                            combinePlot = "none"))
plotDoubletCells <- formatPlotList(plotDoubletCells, samples)

i="DoubletCells"
cat(paste0('## ', i, ' \n'))
```
<br />
`r description_DoubletCells$introduction`
`r description_DoubletCells$runDoubletCells`
`r description_DoubletCells$plotDoubletCellsResults`
`r description_DoubletCells$output`

## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE, results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### DoubletCells Doublet Score {.tabset .tabset-fade} \n\n ")
  print(plotDoubletCells$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  print(plotDoubletCells$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  print(plotDoubletCells$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  DoubletCellMeta <- sce.qc@metadata$runDoubletCell
  # if(length(samples) == 1) {
  #   DoubletCellMeta <- list(DoubletCellMeta)
  #   names(DoubletCellMeta) <- sample
  # }
  
  x <- preprocessMeta(DoubletCellMeta)
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  cat(description_DoubletCells$parameter)
  cat("\n\n")
}
```
`r cat("\n")`




```{r, include=FALSE, warning=FALSE, message=FALSE}
description_CxdsResults<- descriptionCXDS()
```

```{r, "Cxds", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
plotCxds<- suppressWarnings(plotCxdsResults(sce.qc, 
                                            reducedDimName=reduceDimUsed,
                                            sample = colData(sce.qc)[['sample']],
                                            combinePlot = "none"))
plotCxds <- formatPlotList(plotCxds, samples)
i="Cxds"
cat(paste0('## ', i, ' \n'))
```
<br />
`r description_CxdsResults$introduction`
`r description_CxdsResults$runCxds`
`r description_CxdsResults$plotCxdsResults`
`r description_CxdsResults$output`


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE,results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### Cxds Doublet Score {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Cxds Doublet Assignment {.tabset .tabset-fade} \n\n")
  plot(plotCxds$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  CxdsMeta <- sce.qc@metadata$runCxds
  # if(length(samples) == 1) {
  #   CxdsMeta <- list(CxdsMeta)
  #   names(CxdsMeta) <- sample
  # }
  
  x <- preprocessMeta(CxdsMeta)
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  
  cat(description_CxdsResults$nTop)
  cat(description_CxdsResults$binThresh)
  cat(description_CxdsResults$verb)
  cat(description_CxdsResults$retRes)
  cat(description_CxdsResults$estNdbl)
  cat("\n\n")
}
```
`r cat("\n")`





```{r, include=FALSE, warning=FALSE, message=FALSE}
description_BcdsResults<- descriptionBCDS()
```

```{r, "Bcds", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
plotBcds<- suppressWarnings(plotBcdsResults(sce.qc, 
                                           reducedDimName=reduceDimUsed,
                                           sample = colData(sce.qc)[['sample']],
                                           combinePlot = "none"))
plotBcds <- formatPlotList(plotBcds, samples)
i="Bcds"
cat(paste0('## ', i, ' \n'))
```
<br />
`r description_BcdsResults$introduction`
`r description_BcdsResults$runBcds`
`r description_BcdsResults$plotBcdsResults`
`r description_BcdsResults$output`


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE,results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### Bcds Doublet Score {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### Bcds Doublet Assignment {.tabset .tabset-fade} \n\n")
  plot(plotBcds$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  BcdsMeta <- sce.qc@metadata$runBcds
  # if(length(samples) == 1) {
  #   BcdsMeta <- list(BcdsMeta)
  #   names(BcdsMeta) <- sample
  # }
  
  x <- preprocessMeta(BcdsMeta)
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  
  cat(description_BcdsResults$nTop)
  cat(description_BcdsResults$srat)
  cat(description_BcdsResults$nmax)
  cat(description_BcdsResults$varImp)
  cat("\n\n")
}
```
`r cat("\n")`






```{r, include=FALSE, warning=FALSE, message=FALSE}
description_ScdsHybrid<- descriptionScdsHybrid()
```


```{r, "ScdsHybrid", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
plotScdsHybrid<- suppressWarnings(plotScdsHybridResults(sce.qc, 
                                                        reducedDimName=reduceDimUsed,
                                                        sample = colData(sce.qc)[['sample']],
                                                        combinePlot = "none"))
plotScdsHybrid <- formatPlotList(plotScdsHybrid, samples)
i="ScdsHybrid"
cat(paste0('## ', i, ' \n'))
```
<br />
`r description_ScdsHybrid$introduction`
`r description_ScdsHybrid$runCxdsBcdsHybrid`
`r description_ScdsHybrid$plotScdsHybridResults`
`r description_ScdsHybrid$output`


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE,results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### ScdsHybrid Doublet Score {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["scatter_doubletScore"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["density_doubletScore"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["violin_doubletScore"]])
  cat("\n\n")
  
  cat("#### ScdsHybrid Doublet Assignment {.tabset .tabset-fade} \n\n")
  plot(plotScdsHybrid$Sample[[sample]][["scatter_doubletCall"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  HybridMeta <- sce.qc@metadata$runCxdsBcdsHybrid
  # if(length(samples) == 1) {
  #   HybridMeta <- list(HybridMeta)
  #   names(HybridMeta) <- sample
  # }
  
  x <- preprocessMeta(HybridMeta)
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  
  cat(description_ScdsHybrid$parameters)
  cat("\n\n")
}
```
`r cat("\n")`




# Ambient RNA Detection
```{r, include=FALSE, warning=FALSE, message=FALSE}
description_DecontX<- suppressWarnings(descriptionDecontX())
```

```{r, "DecontX", results="asis", fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}
plotDecontX<- suppressWarnings(plotDecontXResults(sce.qc, 
                                                  reducedDimName=reduceDimUsed,
                                                  sample = colData(sce.qc)[['sample']],
                                                  combinePlot = "none"))
plotDecontX <- formatPlotList(plotDecontX, samples)

i="DecontX"
cat(paste0('## ', i, ' \n'))
```
<br />
`r description_DecontX$introduction`
`r description_DecontX$runDecontX`
`r description_DecontX$plotDecontXResults`
`r description_DecontX$output`
`r description_DecontX$contamination`
`r description_DecontX$clustering`


## {.unlisted .unnumbered .toc-ignore .tabset .tabset-fade}
```{r echo=FALSE,results="asis", fig.align="center"}
cat("\n")
for (sample in samples){
  cat(paste0("### ", sample, " {.tabset .tabset-fade} \n"))
  cat("\n\n")
  
  cat("#### DecontX Contamination Score {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["scatter_decontXContamination"]])
  cat("\n\n")
  
  cat("#### Density Score {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["density_decontXContamination"]])
  cat("\n\n")
  
  cat("#### Violin Score {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["violin_decontXContamination"]])
  cat("\n\n")
  
  cat("#### DecontX Clusters {.tabset .tabset-fade} \n\n")
  plot(plotDecontX$Sample[[sample]][["scatter_decontXClusters"]])
  cat("\n\n")
  
  cat("#### Parameters {.tabset .tabset-fade} \n\n")
  DecontXMeta <- sce.qc@metadata$runDecontX
  # if(length(samples) == 1) {
  #   DecontXMeta <- list(DecontXMeta)
  #   names(DecontXMeta) <- sample
  # }
  
  x <- preprocessMeta(DecontXMeta)
  cat(t(as.data.frame(x)) %>%
    knitr::kable(format = "html") %>% kableExtra::kable_styling() %>%
    kableExtra::scroll_box(width = "80%"))
  cat("\n\n")
}
```
`r cat("\n")`

# Session Info
```{r "CellQC-session-info", echo=FALSE}
sessionInfo()
```




