---
title: "Normalizating using Seurat"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params:
  object: object
output:
  html_document:
    number_sections: yes
    toc: yes
    code_folding: hide
---
## Normalization
Before the data can be used in the downstream analysis, normalization techniques that remove technical noise or variation must be applied. For this purpose, Seurat uses a global-scaling normalization method where the input raw data matrix is normalized by dividing the gene expression measurements by the total expression for each cell, multiplies it with a global-scaling factor and log-transforms the result. The output log-transformed and normalized gene expression measurements with technical variation and bias minimized can now better reflect the biological variability in the downstream methods. 

> The input raw data matrix *`r assayNames(data)[1]`* was normalized using a global-scaling normalization method *LogNormalize* that divides the gene expression measurements by the total expression for each cell, multiplies the resultant values with a scaling factor of *10,000* and log-transforms the result. 

```{r Seurat Normalization, message=FALSE, warning=FALSE}
normalizeParams <- list(
  inSCE = data,
  normalizationMethod = "LogNormalize",
  scaleFactor = 10000,
  verbose = FALSE
)
data <- do.call("seuratNormalizeData", normalizeParams)
```

```{r, include=FALSE}
normalizeParams$inSCE <- NULL
metadata(data)$seurat$sctk$report$normalizeParams <- normalizeParams
```
