---
title: "Scale Data using Seurat"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params:
  object: object
output:
  html_document:
    number_sections: yes
    toc: yes
    code_folding: hide
---
## Scaling
Standardizing gene expression values is considered necessary to ensure that dimensionality reduction tools produce unbiased results due to high variance of some features. For this purpose, Seurat performs a *z-score* scaling on the input matrix and standardizes (centers and scales) the expression values of the variable features identified previously before they are input to the downstream methods.  

> The variable features identified previously were scaled using *z-score* scaling method and the output expression values were clipped to maximum and minimum value of 10 & -10 respectively.

```{r Seurat Scaling Data, warning=FALSE}
data <- seuratScaleData(inSCE = data, model = "linear", verbose = FALSE)
```
```{r, include=FALSE}
scaleParams <- list(
  model = "linear", 
  verbose = FALSE)
metadata(data)$seurat$sctk$report$scaleParams <- scaleParams
```
