---
title: "differentialAbundanceFET Results"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params:
  sce: sce
  analysisName: analysisName
  pdf: pdf
  showSession: showSession
output: 
 html_document:
     toc: true
     toc_float: true
     number_sections: true
     theme: cosmo
     code_folding: hide
---

```{r setup, include=FALSE}
require(singleCellTK)
require(celda)
require(knitr)
require(gridExtra)
require(ggplot2)
require(dplyr)
library(kableExtra)

sce <- params$sce
analysisName <- params$analysisName
pdf <- params$pdf
showSession <- params$showSession


dev <- ifelse(isTRUE(pdf), c("png"), c("png", "pdf"))
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  cache.lazy = FALSE,
  # don't do lazy-loading for big objects
  cache.comments = FALSE,
  fig.align = "center",
  fig.keep = "all",
  dev = dev
)

table <- getDiffAbundanceResults(sce, analysisName)

```

# Differential Abundance FET

This report will visualize results from the `diffAbundanceFET` function, which calculates the cell counting and fraction by dividing all cells to groups specified by the arguments, together with statistical summary by performing Fisher Exact Tests (FET). The differential abundance analysis presented here is `r params$analysisName`.

```{r table, echo = FALSE}
table <- table %>% mutate_if(is.numeric, round, digits = 4)
kable(table, style = 'html', row.names = F) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "800px")
```

# Session Information
```{r session, eval = showSession, echo = showSession}
sessionInfo()
```
