---
title: "Introduction to singleCellTK"
author:
- name: David Jenkins
  affiliation: 
  - The Section of Computational Biomedicine, Boston University School of Medicine, Boston, MA
  - Program in Bioinformatics, Boston University, Boston, MA
  email: dfj@bu.edu
- name: Tyler Faits
  affiliation: 
  - The Section of Computational Biomedicine, Boston University School of Medicine, Boston, MA
  - Program in Bioinformatics, Boston University, Boston, MA
- name: W. Evan Johnson
  affiliation: 
  - The Section of Computational Biomedicine, Boston University School of Medicine, Boston, MA
  - Program in Bioinformatics, Boston University, Boston, MA
package: singleCellTK
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Primer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The Single Cell Toolkit (SCTK) is an interactive scRNA-Seq analysis package that
allows a user to upload raw scRNA-Seq count matrices and perform downstream
scRNA-Seq analysis interactively through a web interface, or through a set of
R functions through the command line. The package is written in R with a
graphical user interface (GUI) written in Shiny. Users can perform analysis with
modules for filtering raw results, clustering, batch correction, differential
expression, pathway enrichment, and scRNA-Seq study design, all in a simple to
use point and click interface. The toolkit also supports command line data
processing, and results can be loaded into the GUI for additional exploration
and downstream analysis.

# Installation

> Note: Some package dependencies require Bioconductor v3.6,
> https://bioconductor.org/install/

singleCellTK is under development. You can install the development version from
github:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("compbiomed/singleCellTK")
```

## Troubleshooting Installation

For the majority of users, the commands above will install the latest version of the singleCellTK without any errors. Rarely, you may encounter an error due to previously installed versions of some packages that are required for the singleCellTK. If you encounter an error during installation, use the commands below to check the version of Bioconductor that is installed:

```{r eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocVersion()
```

If the version number is not 3.6 or higher, you must upgrade Bioconductor to install the toolkit:

```{r eval=FALSE}
biocLite("BiocUpgrade")
```

After you install Bioconductor 3.6 or higher, you should be able to install the toolkit using ```devtools::install_github("compbiomed/singleCellTK")```. If you still encounter an error, ensure your Bioconductor packages are up to date by running the following command.

```{r eval=FALSE}
biocValid()
```

If the command above does not return ```TRUE```, run the following command to update your R packages:

```{r eval=FALSE}
biocLite()
```

Then, try to install the toolkit again:

```{r eval=FALSE}
devtools::install_github("compbiomed/singleCellTK")
```

If you still encounter an error, please [contact us](mailto:dfj@bu.edu) and we'd be happy to help.

# Data Structure

The Single Cell Toolkit uses a [SingleCellExperiment](https://www.bioconductor.org/packages/release/bioc/html/SingleCellExperiment.html) object to store data matrices along with annotation information, metadata, and reduced
dimensionality data (PCA, t-SNE, etc.).

# Run the Shiny App

## Example Data

Example data is available within the app. To get started, simply run the singleCellTK function:

```{r eval=FALSE}
library(singleCellTK)
singleCellTK()
```

## Upload data directly through the shiny app

To upload count matrices and annotation information stored as text files, run
the singleCellTK function:

```{r eval=FALSE}
library(singleCellTK)
singleCellTK()
```

Then, follow data upload instructions in the "Upload Tab" vignette.

## Load data from a SingleCellExperiment object

### Creating a SingleCellExperiment object

To create a SingleCellExperiment object, we have provided the ```createSCE()```
function:

```{r}

```

### Loading data stored in a SingleCellExperiment object

# Vignettes

To help you get started with the SCTK, we have prepared several vignettes in two
categories: interactive analysis and R console analysis.

## Interactive Analysis

## R Console Analysis

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
