---
title: "Processing and Visualizing Data in the Single Cell Toolkit"
author:
- name: David Jenkins
  affiliation: 
  - The Section of Computational Biomedicine, Boston University School of Medicine, Boston, MA
  - Program in Bioinformatics, Boston University, Boston, MA
  email: dfj@bu.edu
- name: Tyler Faits
  affiliation: 
  - The Section of Computational Biomedicine, Boston University School of Medicine, Boston, MA
  - Program in Bioinformatics, Boston University, Boston, MA
- name: W. Evan Johnson
  affiliation: 
  - The Section of Computational Biomedicine, Boston University School of Medicine, Boston, MA
  - Program in Bioinformatics, Boston University, Boston, MA
package: singleCellTK
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Primer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The analysis modules available through the Shiny app are also available as R
functions for standard R console processing of single cell RNA-Seq data using
a SCtkExperiment object. At any stage, you can load the Shiny App to
interactively visualize and analyze a data set, but this vignette will show a
standard workflow run entirely through the R console.

# MAITS Example

The MAST package contains a convenient scRNA-Seq example data set of 96 Mucosal
Associated Invariant T cells (MAITs), half of which were stimulated with
cytokines to induce a response. For more details, consult the [MAST package](https://www.bioconductor.org/packages/release/bioc/html/MAST.html)
and [vignette](https://www.bioconductor.org/packages/release/bioc/vignettes/MAST/inst/doc/MAITAnalysis.html).

We will first convert the MAST example dataset to a SCtkExperiment object.

```{r maits_load, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(MAST)
  library(singleCellTK)
  library(xtable)
})

data(maits, package="MAST")
maits_sce <- createSCE(assayFile = t(maits$expressionmat),
                       annotFile = maits$cdat,
                       featureFile = maits$fdat,
                       assayName = "logtpm",
                       inputDataFrames = TRUE,
                       createLogCounts = FALSE)
rm(maits)
```

## summarizeTable

You can get summary metrics with the `summarizeTable` function:

```{r maits_summarize, results='asis'}
print(xtable(summarizeTable(maits_sce, useAssay = "logtpm")),
      type="html", include.rownames=FALSE,
      html.table.attributes='class="table table-condensed"')
```

Typically, these summary statistics would be run on a "counts" matrix, but here
we have log(tpm) values so the average number of reads per cell is calculated
from the normalized values instead of raw counts.

## Filtering by Annotation

Explore the available annotations in the data:

```{r maits_colnames}
colnames(colData(maits_sce))
table(colData(maits_sce)$ourfilter)
```

The data has a filtered dataset with 74 'pass filter' samples, let's subset
the data to include the pass filter samples

```{r maits_filter}
maits_subset <- maits_sce[, colData(maits_sce)$ourfilter]
table(colData(maits_subset)$ourfilter)
```

```{r maits_filter_table, results='asis'}
print(xtable(summarizeTable(maits_subset, useAssay = "logtpm")),
      type="html", include.rownames=FALSE,
      html.table.attributes='class="table table-condensed"')
```

## Visualization

Initially, there are no reduced dimensionality datasets stored in the object

```{r maits_availablereduceddims}
reducedDims(maits_subset)
```

PCA and t-SNE can be added to the object with the getPCA() and getTSNE()
functions:

```{r maits_getpcatsne}
maits_subset <- getPCA(maits_subset, useAssay = "logtpm",
                       reducedDimName = "PCA_logtpm")
maits_subset <- getTSNE(maits_subset, useAssay = "logtpm",
                        reducedDimName = "TSNE_logtpm")
reducedDims(maits_subset)
```

### PCA

PCA data can be visualized with the plotPCA() function:

```{r maits_pca}
plotPCA(maits_subset, reducedDimName = "PCA_logtpm", colorBy = "condition")
```

### t-SNE

t-SNE data can be visualized with the plotTSNE() function:

```{r maits_tsne}
plotTSNE(maits_subset, reducedDimName = "TSNE_logtpm", colorBy = "condition")
```

## Converting Gene Names

The singleCellTK has the ability to convert gene ids to various formats using
the org.*.eg.db Bioconductor annotation packages. These packages are not
installed by default, so these must be manually installed before this function
will work.

```{r maits_convert_symbols, message=FALSE}
suppressPackageStartupMessages({
  library(org.Hs.eg.db)
})
maits_entrez <- maits_subset
maits_subset <- convertGeneIDs(maits_subset, inSymbol = "ENTREZID",
                               outSymbol = "SYMBOL", database = "org.Hs.eg.db")
#to remove confusion for MAST about the gene name:
rowData(maits_subset)$primerid <- NULL
```

## Differential Expression with MAST

MAST is a popular package for performing differential expression analysis on
scRNA-Seq data that models the effect of dropouts using a bimodal distribution
and by including the cellular detection rate into the differential expression
model. Functions in the toolkit allow you to perform this analysis on a
SCtkExperiemnt object.

### Adaptive Thresholding

First, an adaptive threshold is calculated by binning genes with similar
expression levels.

```{r maits_thresh, fig.height=8, message=FALSE}
thresholds <- thresholdGenes(maits_subset, useAssay = "logtpm")
par(mfrow = c(5, 4))
plot(thresholds)
par(mfrow = c(1, 1))
```

### Run MAST

MAST analysis can be run with a single function

```{r maits_MAST, message=FALSE}
mast_results <- MAST(maits_subset, condition = "condition", useThresh = TRUE,
                     useAssay = "logtpm")
```

The resulting significantly differentially expressed genes can be visualized
using a violin plot, linear model, or heatmap:

```{r maits_violin, fig.height=8, message=FALSE}
MASTviolin(maits_subset, useAssay = "logtpm", fcHurdleSig = mast_results,
           threshP = TRUE, variable = "condition")
```

```{r maits_lm, fig.height=8, message=FALSE}
MASTregression(maits_subset, useAssay = "logtpm", fcHurdleSig = mast_results,
               threshP = TRUE, variable = "condition")
```

```{r maits_heatmap}
plotDiffEx(maits_subset, useAssay = "logtpm", condition = "condition",
           geneList = mast_results$Gene[1:100], annotationColors = "auto",
           displayRowLabels = FALSE, displayColumnLabels = FALSE)
```

### Pathway Activity with GSVA

The singleCellTK supports pathway activity analysis using the GSVA package.
Currently, the toolkit supports performing this analysis on human datasets with
entrez IDs. Data can be visualized as a violin plot or a heatmap.

```{r maits_gsva, message=FALSE}
gsvaRes <- gsvaSCE(maits_entrez, useAssay = "logtpm",
                   "MSigDB c2 (Human, Entrez ID only)",
                   c("KEGG_PROTEASOME",
                     "REACTOME_VIF_MEDIATED_DEGRADATION_OF_APOBEC3G",
                     "REACTOME_P53_INDEPENDENT_DNA_DAMAGE_RESPONSE",
                     "BIOCARTA_PROTEASOME_PATHWAY",
                     "REACTOME_METABOLISM_OF_AMINO_ACIDS",
                     "REACTOME_REGULATION_OF_ORNITHINE_DECARBOXYLASE",
                     "REACTOME_CYTOSOLIC_TRNA_AMINOACYLATION",
                     "REACTOME_STABILIZATION_OF_P53",
                     "REACTOME_SCF_BETA_TRCP_MEDIATED_DEGRADATION_OF_EMI1"),
                    parallel.sz=1)

set.seed(1234)
gsvaPlot(maits_subset, gsvaRes, "Violin", "condition")
gsvaPlot(maits_subset, gsvaRes, "Heatmap", "condition")
```

# Batch Effects Example

It is possible to use ComBat within the Single Cell Toolkit. This support is
experimental, since ComBat was not designed for scRNA-Seq. Here, we will load
the bladderbatch example data into a SingleCellExperiment object.

```{r load_bladderbatch, message=FALSE}
library(bladderbatch)
data(bladderdata)
dat <- bladderEset[1:50,]

pheno = pData(dat)
edata = exprs(dat)
bladder_sctke <- createSCE(assayFile = edata,
                           annotFile = pheno,
                           assayName = "microarray",
                           inputDataFrames = TRUE,
                           createLogCounts = FALSE)
assay(bladder_sctke, "combat") <- ComBatSCE(SCEdata = bladder_sctke,
                                            batch = "batch",
                                            useAssay = "microarray",
                                            covariates = "cancer")
```

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
